The cut operation on matrix factorisations

### 2017-07-22 referee-reply.tex

\documentclass[english,letter paper,12pt,leqno]{article}
\usepackage{stmaryrd}
\usepackage{amsmath, amscd, amssymb, mathrsfs, accents, amsfonts,amsthm}
\usepackage[all]{xy}
\usepackage{dsfont}
\usepackage{tikz}
\def\nicedashedcolourscheme{\shadedraw[top color=blue!22, bottom color=blue!22, draw=gray, dashed]}
\def\nicecolourscheme{\shadedraw[top color=blue!22, bottom color=blue!22, draw=white]}
\def\nicepalecolourscheme{\shadedraw[top color=blue!12, bottom color=blue!12, draw=white]}
\def\nicenocolourscheme{\shadedraw[top color=gray!2, bottom color=gray!25, draw=white]}
\def\nicereallynocolourscheme{\shadedraw[top color=white!2, bottom color=white!25, draw=white]}
\definecolor{Myblue}{rgb}{0,0,0.6}
\usepackage[a4paper,colorlinks,citecolor=Myblue,linkcolor=Myblue,urlcolor=Myblue,pdfpagemode=None]{hyperref}

\SelectTips{cm}{}

\setlength{\evensidemargin}{0.1in}
\setlength{\oddsidemargin}{0.1in}
\setlength{\textwidth}{6.3in}
\setlength{\topmargin}{0.0in}
\setlength{\textheight}{8.5in}
\setlength{\headheight}{0in}

\newtheorem{theorem}{Theorem}[section]
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{corollary}[theorem]{Corollary}
\newtheorem{setup}[theorem]{Setup}

\newtheoremstyle{example}{\topsep}{\topsep}
	{}
	{}
	{\bfseries}
	{.}
	{2pt}
	{\thmname{#1}\thmnumber{ #2}\thmnote{ #3}}

	\theoremstyle{example}
	\newtheorem{definition}[theorem]{Definition}
	\newtheorem{example}[theorem]{Example}
	\newtheorem{remark}[theorem]{Remark}
	\newtheorem{strat}[theorem]{Strategy}

\numberwithin{equation}{section}

% Operators
\def\eval{\operatorname{ev}}
\def\res{\operatorname{Res}}
\def\Coker{\operatorname{Coker}}
\def\Ker{\operatorname{Ker}}
\def\im{\operatorname{Im}}
\def\can{\operatorname{can}}
\def\K{\mathbf{K}}
\def\D{\mathbf{D}}
\def\N{\mathbf{N}}
\def\LG{\mathcal{LG}}
\def\Ab{\operatorname{Ab}}
\def\Hom{\operatorname{Hom}}
\def\modd{\operatorname{mod}}
\def\Modd{\operatorname{Mod}}
\def\be{\begin{equation}}
\def\ee{\end{equation}}
\def\nN{\mathds{N}}
\def\nZ{\mathds{Z}}
\def\nQ{\mathds{Q}}
\def\nR{\mathds{R}}
\def\nC{\mathds{C}}
\def\L{\mathcal{C}}
\def\ferm{\gamma}
\def\fermc{\gamma^\dagger}
\DeclareMathOperator{\Ext}{Ext}
\DeclareMathOperator{\Tr}{Tr}
\DeclareMathOperator{\End}{End}
\DeclareMathOperator{\rank}{rank}
\DeclareMathOperator{\tot}{Tot}
\DeclareMathOperator{\ch}{ch}
\DeclareMathOperator{\str}{str}
\DeclareMathOperator{\hmf}{hmf}
\DeclareMathOperator{\HMF}{HMF}
\DeclareMathOperator{\hf}{HF}
\DeclareMathOperator{\At}{At}
\DeclareMathOperator{\Cat}{Cat}
\DeclareMathOperator{\Spec}{Spec}

\begin{document}

% Commands
\def\Res{\res\!}
\newcommand{\ud}{\mathrm{d}}
\newcommand{\Ress}[1]{\res_{#1}\!}
\newcommand{\cat}[1]{\mathcal{#1}}
\newcommand{\lto}{\longrightarrow}
\newcommand{\xlto}[1]{\stackrel{#1}\lto}
\newcommand{\mf}[1]{\mathfrak{#1}}
\newcommand{\md}[1]{\mathscr{#1}}
\def\sus{\l}
\def\l{\,|\,}
\def\sgn{\textup{sgn}}

\title{Reply to referee report}

I thank the referee for their useful comments, and make the following remarks in reply:

\begin{enumerate}
\item The referee is right to observe that while the term ``cut system'' appears prominently in the title and elsewhere in the paper, it is never really defined. After some consideration it seems to me better to remove the term from the paper altogether, and leave the development to sequels where it belongs. To this end I have removed all occurrences of ``cut system'' and referred instead to the ``cut operation'' which (I hope) is clearly defined in the text. Other occurrences of ``cut system'' (for instance in the discussion of linear logic) have been replaced by ``enrichment'' which is a more useful descriptor.

\item p.3 added reference for Atiyah classes.

\item p.5 the logical variable has been renamed to $\alpha$ for clarity.

\item p.8 the $\Psi_a$ are not given a form of $2$-functoriality in $a$, so no.

\item p.10 added a more precise reference

\item p.11 the nonstandard notation has been fixed.

\item p.13 Added $F = k^n$.

\end{enumerate}

In addition I have made some minor typographical corrections.

\end{document}

### 2017-07-22 cutsystems-new.tex
Updates

The bicategory $\LG$ of Landau-Ginzburg models has a rich structure, attested to by the existence of adjoints and pivotality \cite{lgdual}, its role in recent important results in topological field theory \cite{genorb,ade} and the theory of homological knot invariants \cite{kr0401268}. While the objects and $1$-morphisms of $\LG$ are very simple, the composition operation for $1$-morphisms has an ``infinitary'' character. This paper continues the study of this operation begun in \cite{dm1102.2957}.

The objects of $\LG$ are a class of polynomials $W(x)$ called potentials, and $1$-morphisms $W(x) \lto V(y)$ are matrix factorisations of the difference $V(y) - W(x)$ over $\nC[x,y]$. Such a matrix factorisation is a $\nZ_2$-graded free $\nC[x,y]$-module $X = X^0 \oplus X^1$ together with an odd operator $d_X$ satisfying $d_X^2 = (V(y) - W(x)) \cdot 1_X$. These $1$-morphisms form a category denoted $\LG(W,V)$. For a third potential $U(z)$, composition is a functor
\begin{gather*}
\LG(V,U) \times \LG(W,V) \lto \LG(W,U)\,,\\
(Y , X) \mapsto Y \circ X
\end{gather*}
% In this paper and its sequel we study an enrichment of $\LG$ which we refer to as a \emph{cut system}.
defined for matrix factorisations $Y$ of $V - U$ and $X$ of $W - V$ by
\be
Y \circ X = \big( Y \otimes_{\nC[y]} X, d_Y \otimes 1 + 1 \otimes d_X \big)\,.
\ee
This is the analogue of convolution of Fourier-Mukai kernels in the matrix factorisation setting and suffers from the same defect, namely, the resulting matrix factorisation $Y \circ X$ is a free module of \emph{infinite rank} over $\nC[x,z]$, which is only homotopy equivalent to a finite rank matrix factorisation. While this finite model can be proven abstractly to exist \cite{dm1102.2957} there is no obvious way to describe it as an explicit function of the input data $Y,X$. This means that we are stuck with infinite rank matrix factorisations in the theory.

Together with Dyckerhoff we initiated in \cite{dm1102.2957} the study of these finite models in terms of homological perturbation and Atiyah classes, and in the present paper this program reaches its logical conclusion. We introduce a bicategory $\L$ in which the objects are again potentials and the $1$-morphisms $W \lto V$ are finite rank matrix factorisations of $V - W$ additionally equipped with the structure of a representation of a Clifford algebra. This bicategory is equivalent to $\LG$, but it is simpler: the composition operation is described by a finite number of polynomial functions of the input data. More precisely, the composition of $1$-morphisms in $\L$ is a functor
\begin{gather}
\L(V,U) \times \L(W,V) \lto \L(W,U)\,,\label{eq:cut_functor_intro}\\
(Y,X) \mapsto Y \l X
\end{gather}
which we call the \emph{cut operation}. Let us describe the cut in the most important case, where $Y, X$ are matrix factorisations as above, not equipped with any additional structure as representations of a Clifford algebra. The Jacobi algebra $J_V = \nC[y_1,\ldots,y_m]/(\partial_{y_1} V, \ldots, \partial_{y_m} V)$ is a finite-dimensional $\nC$-vector space by the assumption that $V(y)$ is a potential, and the cut is defined to be the following finite rank matrix factorisation over $\nC[x,z]$
\be
Y \l X = Y \otimes_{\nC[y]} J_V \otimes_{\nC[y]} X = ( Y \otimes_{\nC[y]} X ) / (\partial V) (Y \otimes_{\nC[y]} X )
\ee
equipped with a family of odd closed operators $\{\ferm_i, \fermc_i\}_{i=1}^m$ defined by
\begin{equation}\label{eq:intro_clifford_act1_intro}
\ferm_i = \At_i\,, \qquad \fermc_i = - \partial_{y_i}(d_X) - \frac{1}{2} \sum_q \partial_{y_q} \partial_{y_i}(V) \At_{q}
\end{equation}
where the $\At_q$ are Atiyah classes and $\partial_{y_i} = \frac{\partial}{\partial y_i}$. These operators satisfy Clifford relations
\be\label{eq:clifford_relations_intro}
\ferm_i \ferm_j + \ferm_j \ferm_i = 0, \qquad \fermc_i \fermc_j + \fermc_j \fermc_i = 0, \qquad \ferm_i \fermc_j + \fermc_j \ferm_i = \delta_{ij}
\ee
up to homotopy (see Theorem \ref{theorem:cut_is_rep}). The differential on $Y \l X$ and the Clifford operators can be written explicitly as matrices over $\nC[x,z]$, where the coefficients of the monomials in each row and column are polynomials in the monomial coefficients of $d_Y, d_X$ and $V$. Thus, the cut operation is a polynomial function of its inputs and, in particular, infinite rank matrix factorisations do not appear in the definition of $\L$.

Let us sketch where the cut operation comes from, why Clifford algebras are involved, and why there is an equivalence $\L \cong \LG$. In Section \ref{section:intro_appli} we describe various applications. By the results of \cite{dm1102.2957} there is an isomorphism in the homotopy category of (infinite rank) matrix factorisations of $U - W$ over $\nC[x,z]$ of the form
\begin{equation}\label{eq:final_finite_model_intro}
\xymatrix@C+4pc{
Y \l X \ar@<-1ex>[r]_-{\Phi^{-1}} & S_m \otimes_{\nC} ( Y \otimes X )\ar@<-1ex>[l]_-{\Phi}
}
\end{equation}
where $S_m = \bigwedge ( \nC^{\oplus m}[1] )$ denotes the exterior algebra on a $\nZ_2$-graded vector space. The algebra $C_m = \End_{\nC}(S_m)$ is a Clifford algebra. The main theorem of the present paper (Theorem \ref{theorem:htpy_equivalence_main}) identifies the natural action of $C_m$ on the right hand side of \eqref{eq:final_finite_model_intro} with the action of $C_m$ via the aforementioned operators $\ferm_i, \fermc_i$ on the left hand side.

There is a category called the \emph{Clifford thickening} of the homotopy category of matrix factorisations of $U - W$, in which objects are matrix factorisations equipped with a $C_l$-action for some $l$ (i.e. $l$ is allowed to be different for different objects). In this category $S_m \otimes_{\nC} ( Y \otimes X )$ with its $C_m$-action is isomorphic to $Y \otimes X$ with no Clifford action, and so \eqref{eq:final_finite_model_intro} can be read as a natural isomorphism $Y \l X \cong Y \otimes X$. We prove this is suitably functorial in $Y,X$ and therefore defines an equivalence $\L \cong \LG$.

@@

The paper is structured as follows: in Section \ref{section:background} we recall the definition of superbicategories and Clifford algebras, and introduce the Clifford thickening of a supercategory. The reason we need the formalism of supercategories is that we make extensive use of Clifford actions on objects of linear categories, and for this to make sense the categories themselves must be $\nZ_2$-graded (that is, supercategories). In Section \ref{section:lg_cut_system} we define $\L$ and in Section \ref{section:equivllg} we prove it is equivalent to $\LG$ as a superbicategory without units.

@@

As a special case of the deformation retract \eqref{eq:final_finite_model_intro} we obtain in Section \ref{example:computing_homs} for any potential $V(y)$ and matrix factorisations $X,X'$ of $V$ a deformation retract of $\nZ_2$-graded complexes of vector spaces
\begin{equation}\label{eq:final_finite_model_hom}
\xymatrix@C+4pc{
(X')^{\vee} \l X \ar@<-1ex>[r]_-{\Phi^{-1}} & S_m \otimes_{\nC} \Hom_{\nC[y]}(X', X) \ar@<-1ex>[l]_-{\Phi}
}
\end{equation}
where $\Hom_{\nC[y]}(X',X)$ denotes the Hom-complex. Observe that this is infinite-dimensional (with finite-dimensional cohomology) while $(X')^{\vee} \l X$ is finite-dimensional.

In the case $X = X'$ the right hand side is a DG-algebra and the deformation retract may be taken as the input to the standard algorithms for producing an $A_\infty$-minimal model of $\End_{\nC[y]}(X)$. It turns out that this input is well-suited for actually doing computations, which we hope to return to elsewhere.

\subsubsection{A geometric enrichment of $\LG$}

Our primary motivation for introducing $\cat{C}$ is that it can be enriched over affine schemes; this is not true directly of $\LG$, because the composition is infinitary. For context we include a sketch. Given potentials $W(x), V(y)$, a matrix factorisation of $V(y) - W(x)$ is after choosing a homogeneous basis just a matrix
\be\label{eq:intro_capitalD}
D \in M_{2r}( \nC[x,y] ) \qquad \text{ satisfying } \qquad D^2 = (V(y) - W(x)) \cdot I_{2r}
\ee
where we require that, dividing $D$ into $r \times r$ blocks, the upper left and bottom right blocks are zero matrices. If we fix the size $r$ of the blocks and bound the degrees of the monomials appearing in $D$ by some integer $k$, we can parametrise matrix factorisations by the $\nC$-points of an affine scheme whose coordinates are the coefficients of the various monomials in each row and column of $D$.

We may add further coordinates and equations to encode closed odd operators satisfying Clifford relations \eqref{eq:clifford_relations_intro} and, in this way, for a choice of parameters $\lambda = (r, k, \ldots)$ define a scheme $\mathscr{M}_{\lambda}(W,V)$ such that as sets
\begin{equation}\label{eq:param_frmf}
\cat{C}(W,V) = \bigcup_{\lambda} \mathscr{M}_{\lambda}(W,V)_{\nC}
\end{equation}
where $Z_{\nC}$ denotes the $\nC$-points of a scheme $Z$. The cut operation is a polynomial function of its inputs, so there is an indexed family of morphisms of schemes
\be\label{eq:enrichment_morphisms_intro}
\mathscr{M}_\mu(V,U) \times \mathscr{M}_\lambda(W,V) \lto \mathscr{M}_{f(\lambda, \mu)}(W, U)
\ee
lifting the functor \eqref{eq:cut_functor_intro} on the level of objects for some function $f(\lambda, \mu)$ of indices. Finally, using the approach of Section \ref{example:computing_homs}, we can define for each pair of potentials $W,V$ and tuple of parameters $\lambda$ a bundle of Clifford representations
\be\label{eq:enrichment_intro_bundles}
\xymatrix{
\mathscr{H}\! om_{W,V,\lambda} \ar[d]\\
\mathscr{M}_\lambda(W,V) \times \mathscr{M}_\lambda(W,V)\,.
}
\ee
Given two finite-rank matrix factorisations $X,X'$ of $V - W$ the fiber of this vector bundle over the point $(X',X)$ is the Clifford representation $(X')^{\vee} \l X$ of \eqref{eq:final_finite_model_hom}, which is isomorphic to the Hom-complex $\Hom_{\nC[x,y]}(X',X)$ tensored with a suitable spinor representation. The cut system which enriches $\cat{C}$ consists of the family of schemes $\mathscr{M}_\lambda(W,V)$, the composition morphisms \eqref{eq:enrichment_morphisms_intro}, the bundles \eqref{eq:enrichment_intro_bundles}, and various other data satisfying natural constraints which encode the structure of the bicategory.

@@

In forthcoming work we use the above enrichment of $\L$ to define a semantics of linear logic \cite{girard_llogic} in which formulas are interpreted by tuples $(\md{X}, \nC[x], W)$ where $\md{X}$ is a scheme and $W \in \nC[x]$ is a potential. Proofs in the logic are interpreted by certain correspondences between these pairs. Let us briefly explain the general point which leads us to use $\L$ rather than $\LG$, for which we will use the denotation of the Church numeral $\underline{2}$ (see \cite{murfet_ll}).

Let $x$ be a variable of the logic and suppose it has denotation $\llbracket x \rrbracket = (\Spec(\nC), \nC[x], W)$ in the semantics. Let $\mathscr{M} = \mathscr{M}(W,W)$ be the (indexed) affine scheme parametrising $1$-morphisms $X: W \lto W$ in $\L$, as discussed above. Then the semantics works as follows:
\begin{gather*}
\llbracket x \multimap x \rrbracket = \big(\Spec(\nC), \nC[x,x'], W(x) - W(x') \big)\,,\\
\llbracket {!}( x \multimap x ) \rrbracket = \big( \mathscr{M}, \nC, 0 \big)\,.
\end{gather*}
A proof of the sequent ${!}(x \multimap x) \vdash x \multimap x$ has for its denotation a correspondence between $\llbracket {!}( x \multimap x ) \rrbracket$ and $\llbracket x \multimap x \rrbracket$, which is a matrix factorisation of $W(x) - W(x')$ over the scheme $\mathscr{M} \times \Spec(\nC[x,x'])$. For example, the Church numeral $\underline{2}$ has for its denotation the matrix factorisation whose fiber over a loop $X: W \lto W$, viewed as a point of $\mathscr{M}$, is the square $X \l X$ as a matrix factorisation of $W(x) - W(x')$ over $\Spec(\nC[x,x'])$. For these definitions to work, the coefficients of the monomials in the matrix factorisation $X \l X$ need to be given explicitly as polynomial functions of the coefficients in $X$. This means we have to use $\L$ rather than $\LG$.

@@

Throughout $k$ is a commutative $\mathbb{Q}$-algebra. By default categories and functors are $k$-linear.

@@

\begin{definition} Let $A$ be a $\mathbb{Z}_2$-graded $k$-algebra and $\cat{C}$ a supercategory. A \emph{left $A$-module in $\cat{C}$} is an object $X \in \cat{C}$ together with a morphism of $\mathbb{Z}_2$-graded algebras $A \lto \cat{C}^*(X,X)$. A \emph{morphism} of left $A$-modules is a morphism in $\cat{C}$ which commutes with the $A$-action in the obvious sense. The category of left $A$-modules in $\cat{C}$ is denoted $\cat{C}_A$. Right modules are defined similarly, using $A^{\operatorname{op}}$.
\end{definition}

@@

The bicategory of Landau-Ginzburg models $\LG$ has for its objects pairs $(x,W)$ consisting of an ordered set of variables $x = (x_1,\ldots,x_n)$ and a potential $W \in k[x] = k[x_1,\ldots,x_n]$. Given potentials $W \in k[x]$ and $V \in k[y]$ the supercategory $\cat{LG}(W,V)$ is
\[
\cat{LG}(W,V) = \hmf(k[x,y], V - W)^{\oplus}\,.
\]
Given a third potential $U \in k[z]$ the composition law is a functor
\[
\LG(W,V) \otimes_k \LG(V, U) \lto \LG(W, U)
\]
which sends a pair of $1$-morphisms $X: W \lto V$ and $Y: V \lto U$ to the tensor product
\begin{equation}\label{eq:tensor_comp}
Y \otimes X = ( Y \otimes_{k[y]} X, d_{Y \otimes X} = d_Y \otimes 1 + 1 \otimes d_X )\,.
\end{equation}
Thus $Y \circ X := Y \otimes X$. This statement requires some care: $k[x,y,z]$ is an infinite rank free module over $k[x,z]$ and so $Y \otimes X$ is an infinite rank matrix factorisation of $U - W$ over $k[x,z]$. However one can prove that $Y \otimes X$ is a direct summand of a finite rank matrix factorisation \cite{dm1102.2957} and therefore a valid object in $\LG(W,U)$.

>>>>

The bicategory $\LG_k$ of Landau-Ginzburg models \cite{lgdual} over a noetherian $\mathbb{Q}$-algebra $k$ is a pivotal bicategory defined in terms of the geometry of isolated hypersurface singularities. This bicategory has various applications \cite{genorb,ade, kr0401268} but an obstacle to further development of the theory is that, while the objects, $1$-morphisms and $2$-morphisms of $\LG_k$ are simply described, the composition operation for $1$-morphisms has an ``infinitary'' character which makes it hard to work with. The current paper aims to resolve this problem. % This paper continues the study of this operation begun in \cite{dm1102.2957}.

The objects of $\LG_k$ are a class of polynomials $W(x)$ called potentials, and $1$-morphisms $W(x) \lto V(y)$ are finite-rank matrix factorisations \cite{eisenbud} of the difference $V(y) - W(x)$ over $k[x,y]$. The idempotent completion of the homotopy category of such matrix factorisations is denoted $\LG_k(W,V)$. For a third potential $U(z)$, composition in $\LG_k$ is a functor
\begin{gather*}
\LG_k(V,U) \times \LG_k(W,V) \lto \LG_k(W,U)\,,\\
(Y , X) \longmapsto Y \circ X
\end{gather*}
% In this paper and its sequel we study an enrichment of $\LG_k$ which we refer to as a \emph{cut system}.
defined for matrix factorisations $Y$ of $V - U$ and $X$ of $W - V$ by
\be
Y \circ X = \big( Y \otimes_{k[y]} X, d_Y \otimes 1 + 1 \otimes d_X \big)\,.
\ee
This is the analogue of convolution of Fourier-Mukai kernels in the matrix factorisation setting and suffers from the same defect, namely, the resulting matrix factorisation $Y \circ X$ is a free module of \emph{infinite rank} over $k[x,z]$. The problem of describing this \emph{a priori} infinite object in some finite way was studied in \cite{dm1102.2957} and the partial solution provided there was to explicitly describe a finite rank matrix factorisation together with an idempotent, splitting in the homotopy category of matrix factorisations to $Y \circ X$. However this solution is not sufficiently functorial to extent to a coherent description of the entire bicategory $\LG_k$.

The current paper refines \cite{dm1102.2957} by presenting the idempotents developed there as arising from representations of Clifford algebras (see Remark \ref{remark:relation_to_toby_paper}). This leads us to define a new bicategory $\L$ in which the objects are again potentials and the $1$-morphisms $W \lto V$ are finite rank matrix factorisations of $V - W$ additionally equipped with the structure of a representation of a Clifford algebra. This bicategory is equivalent to $\LG_k$, but it is simpler: the composition operation is described by a finite number of polynomial functions of the input data. More precisely, the composition of $1$-morphisms in $\L$ is a functor
\begin{gather}
\L(V,U) \times \L(W,V) \lto \L(W,U)\,,\label{eq:cut_functor_intro}\\
(Y,X) \longmapsto Y \l X
\end{gather}
which we call the \emph{cut operation}, and this operation is polynomial in the sense that the coefficients in the polynomials making up the matrix factorisation $Y \l X$ and its Clifford action are themselves polynomial functions of the coefficients defining $X,Y$.

 Let us describe the cut in the most important case, where $Y, X$ are matrix factorisations as above, not equipped with any additional structure as representations of a Clifford algebra. The Jacobi algebra $J_V = k[y_1,\ldots,y_m]/(\partial_{y_1} V, \ldots, \partial_{y_m} V)$ is a finitely generated free $k$-module by the assumption that $V(y)$ is a potential, and the cut is defined to be the following finite rank matrix factorisation over $k[x,z]$
\be
Y \l X = Y \otimes_{k[y]} J_V \otimes_{k[y]} X
\ee
equipped with a family of odd closed $k[x,z]$-linear operators $\{\ferm_i, \fermc_i\}_{i=1}^m$ defined by
\begin{equation}\label{eq:intro_clifford_act1_intro}
\ferm_i = \At_i\,, \qquad \fermc_i = - \partial_{y_i}(d_X) - \frac{1}{2} \sum_q \partial_{y_q} \partial_{y_i}(V) \At_{q}
\end{equation}
where the $\At_q$ are Atiyah classes (Definition \ref{defn:atiyah}). These operators satisfy Clifford relations
\be\label{eq:clifford_relations_intro}
\ferm_i \ferm_j + \ferm_j \ferm_i = 0, \qquad \fermc_i \fermc_j + \fermc_j \fermc_i = 0, \qquad \ferm_i \fermc_j + \fermc_j \ferm_i = \delta_{ij}
\ee
up to homotopy (see Theorem \ref{theorem:cut_is_rep}). The differential on $Y \l X$ and the Clifford operators can be written explicitly as matrices over $k[x,z]$, where the coefficients of the monomials in each row and column are polynomials in the monomial coefficients of $d_Y, d_X$ and $V$. Thus, the cut operation is a polynomial function of its inputs and, in particular, infinite rank matrix factorisations do not appear in the definition of $\L$.

Let us sketch where the cut operation comes from, why Clifford algebras are involved, and why there is an equivalence $\L \cong \LG_k$. In Section \ref{section:intro_appli} we describe various applications. By the results of \cite{dm1102.2957} there is an isomorphism in the homotopy category of (infinite rank) matrix factorisations of $U - W$ over $k[x,z]$ of the form
\begin{equation}\label{eq:final_finite_model_intro}
\xymatrix@C+4pc{
Y \l X \ar@<-1ex>[r]_-{\Phi^{-1}} & S_m \otimes_{k} ( Y \otimes X )\ar@<-1ex>[l]_-{\Phi}
}
\end{equation}
where $S_m = \bigwedge ( k^{\oplus m}[1] )$ denotes the exterior algebra as a $\nZ_2$-graded module, on the space $k^{\oplus m}$ placed in degree one. The main theorem (Theorem \ref{theorem:htpy_equivalence_main}) identifies the natural action of the Clifford algebra $C_m = \End_{k}(S_m)$ on the right hand side of \eqref{eq:final_finite_model_intro} with the action of $C_m$ via the aforementioned operators $\ferm_i, \fermc_i$ on the left hand side.

There is a category called the \emph{Clifford thickening} of the homotopy category of matrix factorisations of $U - W$, in which objects are matrix factorisations equipped with a $C_l$-action for some $l$ (i.e. $l$ is allowed to be different for different objects). In this category $S_m \otimes_{k} ( Y \otimes X )$ with its $C_m$-action is isomorphic to $Y \otimes X$ with no Clifford action, and so \eqref{eq:final_finite_model_intro} can be read as a natural isomorphism $Y \l X \cong Y \otimes X$. We prove this is suitably functorial in $Y,X$ and therefore defines an equivalence $\L \cong \LG_k$.

@@

The paper is structured as follows: in Section \ref{section:background} we recall the definition of superbicategories and Clifford algebras, and introduce the Clifford thickening of a supercategory. The reason we need the formalism of supercategories is that we make extensive use of Clifford actions on objects of linear categories, and for this to make sense the categories themselves must be $\nZ_2$-graded (that is, supercategories). In Section \ref{section:lg_cut_system} we define $\L$ and in Section \ref{section:equivllg} we prove it is equivalent to $\LG_k$ as a superbicategory without units.

@@

As a special case of the deformation retract \eqref{eq:final_finite_model_intro} we obtain in Section \ref{example:computing_homs} for any potential $V(y)$ and matrix factorisations $X,X'$ of $V$ a deformation retract of $\nZ_2$-graded complexes of $k$-modules
\begin{equation}\label{eq:final_finite_model_hom}
\xymatrix@C+4pc{
(X')^{\vee} \l X \ar@<-1ex>[r]_-{\Phi^{-1}} & S_m \otimes_{k} \Hom_{k[y]}(X', X) \ar@<-1ex>[l]_-{\Phi}
}
\end{equation}
where $\Hom_{k[y]}(X',X)$ denotes the Hom-complex. Observe that if $k$ is a field this is infinite-dimensional (with finite-dimensional cohomology) while $(X')^{\vee} \l X$ is finite-dimensional.

In the case $X = X'$ the right hand side is a DG-algebra and the deformation retract may be taken as the input to the standard algorithms for producing an $A_\infty$-minimal model of $\End_{k[y]}(X)$. It turns out that this input is well-suited for actually doing computations, which we hope to return to elsewhere.

\subsubsection{A geometric enrichment of $\LG_k$}

Our primary motivation for introducing $\cat{C}$ is that it can be enriched over affine schemes; this is not true directly of $\LG_k$, because the composition is infinitary. For context we include a sketch, assuming $k$ is an algebraically closed field of characteristic zero. Given potentials $W(x), V(y)$, a matrix factorisation of $V(y) - W(x)$ is after choosing a homogeneous basis just a matrix
\be\label{eq:intro_capitalD}
D \in M_{2r}( k[x,y] ) \qquad \text{ satisfying } \qquad D^2 = (V(y) - W(x)) \cdot I_{2r}
\ee
where we require that, dividing $D$ into $r \times r$ blocks, the upper left and bottom right blocks are zero matrices. If we fix the size $r$ of the blocks and bound the degrees of the monomials appearing in $D$ by some integer $k$, we can parametrise matrix factorisations by the $k$-points of an affine scheme whose coordinates are the coefficients of the various monomials in each row and column of $D$.

We may add further coordinates and equations to encode closed odd operators satisfying Clifford relations \eqref{eq:clifford_relations_intro} and, in this way, for a choice of parameters $\lambda = (r, k, \ldots)$ define a scheme $\mathscr{M}_{\lambda}(W,V)$ such that as sets
\begin{equation}\label{eq:param_frmf}
\cat{C}(W,V) = \bigcup_{\lambda} \mathscr{M}_{\lambda}(W,V)_{k}
\end{equation}
where $Z_{k}$ denotes the $k$-points of a scheme $Z$. The cut operation is a polynomial function of its inputs, so there is an indexed family of morphisms of schemes
\be\label{eq:enrichment_morphisms_intro}
\mathscr{M}_\mu(V,U) \times \mathscr{M}_\lambda(W,V) \lto \mathscr{M}_{f(\lambda, \mu)}(W, U)
\ee
lifting the functor \eqref{eq:cut_functor_intro} on the level of objects for some function $f(\lambda, \mu)$ of indices. Finally, using the approach of Section \ref{example:computing_homs}, we can define for each pair of potentials $W,V$ and tuple of parameters $\lambda$ a bundle of Clifford representations
\be\label{eq:enrichment_intro_bundles}
\xymatrix{
\mathscr{H}\! om_{W,V,\lambda} \ar[d]\\
\mathscr{M}_\lambda(W,V) \times \mathscr{M}_\lambda(W,V)\,.
}
\ee
Given two finite-rank matrix factorisations $X,X'$ of $V - W$ the fiber of this vector bundle over the point $(X',X)$ is the Clifford representation $(X')^{\vee} \l X$ of \eqref{eq:final_finite_model_hom}, which is isomorphic to the Hom-complex $\Hom_{k[x,y]}(X',X)$ tensored with a suitable spinor representation. The enrichment of $\cat{C}$ consists of the family of schemes $\mathscr{M}_\lambda(W,V)$, the composition morphisms \eqref{eq:enrichment_morphisms_intro}, the bundles \eqref{eq:enrichment_intro_bundles}, and various other data satisfying natural constraints which encode the structure of the bicategory.

@@

In forthcoming work we use the above enrichment of $\L$ to define a semantics of linear logic \cite{girard_llogic} in which formulas are interpreted by tuples $(\md{X}, k[x], W)$ where $\md{X}$ is a scheme and $W \in k[x]$ is a potential. Proofs in the logic are interpreted by certain correspondences between these pairs. Let us briefly explain the general point which leads us to use $\L$ rather than $\LG$, for which we will use the denotation of the Church numeral $\underline{2}$ (see \cite{murfet_ll}).

Let $\alpha$ be a variable of the logic and suppose it has denotation $\llbracket \alpha \rrbracket = (\Spec(k), k[x], W)$ in the semantics. Let $\mathscr{M} = \mathscr{M}(W,W)$ be the (indexed) affine scheme parametrising $1$-morphisms $X: W \lto W$ in $\L$, as discussed above. Then the semantics works as follows:
\begin{gather*}
\llbracket \alpha \multimap \alpha \rrbracket = \big(\Spec(k), k[x,x'], W(x) - W(x') \big)\,,\\
\llbracket {!}( \alpha \multimap \alpha ) \rrbracket = \big( \mathscr{M}, k, 0 \big)\,.
\end{gather*}
A proof of the sequent ${!}(\alpha \multimap \alpha) \vdash \alpha \multimap \alpha$ has for its denotation a correspondence between $\llbracket {!}( \alpha \multimap \alpha ) \rrbracket$ and $\llbracket \alpha \multimap \alpha \rrbracket$, which is a matrix factorisation of $W(x) - W(x')$ over the scheme $\mathscr{M} \times \Spec(k[x,x'])$. For example, the Church numeral $\underline{2}$ has for its denotation the matrix factorisation whose fiber over a loop $X: W \lto W$, viewed as a point of $\mathscr{M}$, is the square $X \l X$ as a matrix factorisation of $W(x) - W(x')$ over $\Spec(k[x,x'])$. For these definitions to work, the coefficients of the monomials in the matrix factorisation $X \l X$ need to be given explicitly as polynomial functions of the coefficients in $X$. This means we have to use $\L$ rather than $\LG_k$.

@@

Throughout $k$ is a noetherian $\mathbb{Q}$-algebra.\footnote{The noetherian hypothesis is only needed in Section \ref{section:superbicatLG} when we pass from a ring to its adic completion with respect to the Jacobi ideal, and we wish to know the morphism from the ring to its completion is flat. This completion is only done to ensure the existence of a suitable connection - if a connection can be provided by other means (e.g. a grading) then the noetherian hypothesis could be dropped.} By default categories and functors are $k$-linear. We write $\partial_x$ for the differential operator $\frac{\partial}{\partial x}$.

@@

\begin{definition}\label{defn:algebra_modules} Let $A$ be a $\mathbb{Z}_2$-graded $k$-algebra and $\cat{C}$ a supercategory. A \emph{left $A$-module in $\cat{C}$} is an object $X \in \cat{C}$ together with a morphism of $\mathbb{Z}_2$-graded algebras $A \lto \cat{C}^*(X,X)$. A \emph{morphism} of left $A$-modules is a morphism in $\cat{C}$ which commutes with the $A$-action in the obvious sense. The category of left $A$-modules in $\cat{C}$ is denoted $\cat{C}_A$. Right modules are defined similarly, using $A^{\operatorname{op}}$.
\end{definition}

@@

The bicategory of Landau-Ginzburg models $\LG = \LG_k$ has for its objects pairs $(x,W)$ consisting of an ordered set of variables $x = (x_1,\ldots,x_n)$ and a potential $W \in k[x] = k[x_1,\ldots,x_n]$. Given potentials $W \in k[x]$ and $V \in k[y]$ the supercategory $\cat{LG}(W,V)$ is
\[
\cat{LG}(W,V) = \hmf(k[x,y], V - W)^{\oplus}\,.
\]
Given a third potential $U \in k[z]$ the composition law is a functor
\[
\LG(W,V) \otimes_k \LG(V, U) \lto \LG(W, U)
\]
which sends a pair of $1$-morphisms $X: W \lto V$ and $Y: V \lto U$ to the tensor product
\begin{equation}\label{eq:tensor_comp}
Y \otimes X = ( Y \otimes_{k[y]} X, d_{Y \otimes X} = d_Y \otimes 1 + 1 \otimes d_X )\,.
\end{equation}
Thus $Y \circ X := Y \otimes X$. This statement requires some care: $k[x,y,z]$ is an infinite rank free module over $k[x,z]$ and so $Y \otimes X$ is an infinite rank matrix factorisation of $U - W$ over $k[x,z]$. However one can prove that $Y \otimes X$ is a direct summand of a finite rank matrix factorisation \cite{dm1102.2957} and therefore a valid object in $\LG(W,U)$.
